A Kochen-Specker system has at least 21 vectors\\ {\small (extended abstract)}

### 2014-09-12 paper/main.tex
paper: more suggestions of Sander

Consider the following experiment. Measure the SPIN of a spin 1
particle along a direction. Or more concretely: shoot a deuterium
atom through a certain fixed inhomogeneous magnetic field. Depending
on the direction it was shot through the field, the particle will
move undisturbed or deviate.

Quantum Mechanics only predicts the probability, given the direction,
whether the particle will deviate.
Its probabilistic prediction has been thoroughly tested.
One wonders: is there a deterministic theory predicting the
outcome of this experiment?

Kochen and Specker have shown that such a deterministic theory must
be odd: it cannot satisfy the plausible SPIN axiom, that is:
\begin{spin}
    Given three pairwise orthogonal directions.
    In exactly one of the directions, the particle will not deviate.
\end{spin}
Their argument is based on the existence of a Kochen-Specker system.
\begin{dfn}
    A \keyword{Kochen-Specker (KS) system} is
    a finite set of points on the sphere\footnote{
            We define KS systems to be three dimensional,
            as in the original proof of Kochen and Specker.
            Later, higher dimensional systems have been studied.
            See, for instance~\cite[p.~201]{qtcm}.
        }
    for which each pair is not antipodal and
    there is no~\keyword{010-coloring}.
    A $010$-coloring is a~$\{0,1\}$-coloring of the points such that%
        \footnote{
                In other papers, like \cite{aow11},
                the~$0$ and~$1$ are swapped; they consider 101-colorings.
                These colorings are of course equivalent and the
                difference arises from considering either squared
                spin measurements $S^2$, or $1-S^2$.
                %Then the colors correspond directly to the squared measurement
               	%outcome of the spin.
               	}
    \begin{enumerate}
        \item
            no pair of orthogonal points are both colored~$1$ and
        \item
            of three pairwise orthogonal points exactly one is colored~$1$;
            or alternatively: they are colored~$0$, $1$ and~$0$ in some order.
    \end{enumerate}
\end{dfn}
% TODO shorten this proof.
A point on the sphere obviously corresponds to a direction in space.
Because of this, the term point, vector and direction
can be used interchangeably. Antipodal points correspond to opposite
vectors and these span the same direction in space.

>>>>

Consider the following experiment. Measure the spin of a spin 1
particle along a direction. For example, shoot a deuterium
atom through a certain fixed inhomogeneous magnetic field,
such as that in the Stern-Gerlach experiment. Depending
on the direction it was shot through the field, the particle will
move undisturbed or deviate.

Quantum Mechanics only predicts the probability, given the direction,
whether the particle will deviate.
Its probabilistic prediction has been thoroughly tested.
One wonders: is there a
\emph{deterministic} theory predicting the
outcome of this experiment?

Kochen and Specker have shown that such a non-contextual deterministic
theory must be odd: it cannot satisfy the plausible SPIN
axiom, that is:
\begin{spin}
    Given three pairwise orthogonal directions.
    In exactly one of the directions, the particle will not deviate.
\end{spin}
Their argument is based on the existence of a Kochen-Specker system.
\begin{dfn}
    A \keyword{Kochen-Specker (KS) system} is
    a finite set of points on the sphere\footnote{
            We define KS systems to be three dimensional,
            as in the original proof of Kochen and Specker.
            Later, higher dimensional systems have been studied.
            See, for instance~\cite[p.~201]{qtcm}.
        }
    for which each pair is not antipodal and
    there is no~\keyword{010-coloring}.
    A $010$-coloring is a~$\{0,1\}$-coloring of the points such that%
        \footnote{
                In other papers, like \cite{aow11},
                the~$0$ and~$1$ are swapped; they consider 101-colorings.
                These colorings are of course equivalent and the
                difference arises from considering either squared
                spin measurements $S^2$, or $1-S^2$.
                %Then the colors correspond directly to the squared measurement
               	%outcome of the spin.
               	}
    \begin{enumerate}
        \item
            no pair of orthogonal points are both colored~$1$ and
        \item
            of three pairwise orthogonal points exactly one is colored~$1$;
            or alternatively: they are colored~$0$, $1$ and~$0$ in some order.
    \end{enumerate}
\end{dfn}
% TODO shorten this proof.
A point on the sphere obviously corresponds to a direction in space.
Because of this, the term point, vector and direction
can be used interchangeably. Antipodal points correspond to opposite
vectors and these span the same direction in space.
